---
title: "HR_APIJET_RouteShiftedToTrajectoryAligned_StateDeviations"
author: "Hari Rallapalli"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---
# Introduction and setting the stage with a logic fault

Presented below are further revised analyses of the KCLKMEM dataset. Please bear in mind that these are still only Ownship-derived data. Nothing from TAP-suggested re-route yet. PLEASE NOTE: I think I found an error in my flightplan identification logic. Currently I do the following search:

1. For each Route index, make a {set} of {waypoint identifiers} (e.g. BOOYA, my favorite) and {altitudes}
2. Compare these sets across sequential Route indicies (e.g. RouteID 1 vs RouteID 2)
3. If there is a difference, and {waypoint identifiers} and {altitudes} from RouteID 2 are a subset of RouteID 1, then this is the same route
4. If there is a difference, and there is no valid subset, then this is a different route.

This logic works ok if there is an added waypoint, modified waypoint, or multiple waypoint reroutes. However, it will fail to detect a change if ATC drops a waypoint from the route!!! The new route will stil be a subset of the old route, but is in fact a new route. I need to update this logic to account for this case. I have an idea for how to account for this. AopRouteDataRecord has a header column called "RteChgType" which has the values "ACT_PILOT_ENTRY_EXEC" or "UNKNOWN". I make the assumption that this is actually a flag for when a route is change (i.e. Pilot is executing an entry). If true, I can us this flag in my logic to account for the route changes. Will do this for the next version of analysis scripts.

Once corrected, I can pass over all of the FDX data with these analysis scripts programatically.

```{r, warning=FALSE}
suppressMessages({
library(tidyverse)
library(ggplot2)
library(maps)
library(multcomp)
})
theme_set(
  theme_minimal()
  )
select <- dplyr::select
```

## Data importing

Updated the Route alignment to Trajectory logic to only start after the first Trajectory is printed. This avoids "missing" Route changes before we can estimate behavior (I counted a 55 second lag between the first Route data block and first Trajectory block being printed). This is now sorted, but I need still need to revise the flightplan identifier logic...
```{r}
 OwnshipTrajectoryFrame <- read.csv("C:\\Users\\Hari.rallapalli\\Desktop\\APIJET\\ROUTEALIGNED_AopOwnshipTrajectoryDataRecord.csv", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(csvsource = "Trajectory")
```

```{r}
OwnshipStateFrame <- read.csv("C:\\Users\\Hari.rallapalli\\Desktop\\APIJET\\DEVIATIONS_AopOwnshipStateDataRecord.csv", header = TRUE, stringsAsFactors = FALSE) %>%
  mutate(csvsource = "State")
```

## Route plotting
Changed the colors a little because I looked at Digital Winglets UI. It's hard to match colors exactly without getting a .svg (or similar) of the map, but I did what I could.
```{r}
usa_map <- map_data("usa")
flightmap <- ggplot(usa_map, aes(x = long, y = lat)) + 
geom_polygon(aes(group=group), fill="gray75") +
 geom_line(data = OwnshipTrajectoryFrame, aes(x = Longitude, y = Latitude, color = as.factor(TrajId)), size = 1) +
  scale_color_brewer("TrajID", palette = "PRGn") +
  geom_point(data = OwnshipStateFrame, aes(x = LongitudePresent, y = LatitudePresent), size = 0.2, color = "cyan") 

flightmap
```

Route prediction is much improved after updating to TrajectoryID = 3, and will improve further once I correct the logic. See below for visual example of the problem with the logic. Cyan line (state) goes directly past a waypoint that is still visible in the purple line (trajectory)

```{r}
usa_map <- map_data("usa")
flightmap_path <- ggplot()+
 geom_line(data = OwnshipTrajectoryFrame, aes(x = Longitude, y = Latitude), size = 1, color = "purple") +
  geom_point(data = OwnshipStateFrame, aes(x = LongitudePresent, y = LatitudePresent), size = 0.2, color = "cyan") +
  xlim(c(-90,-82.5))+
  ylim(c(35, 40))

flightmap_path
```



## State deviation from Trajectory

As we discussed last week, I am separating lateral deviation from altitude deviation (lat/long vs altitude). After separation, it is apparent that the dominant source of deviation from the overall distance trends I showed last week is lateral deviation. The magnitude of altitude deviations (0 - 161 ft) is not within 2 orders of magnitude of lat/long deviations (0 - >52,000 ft). I think this is to be expected.

```{r}
latlong_deviation_plot <- ggplot(data = OwnshipStateFrame, aes(x = time, y = deviation)) +
  geom_line(aes(x = Timestamp, y = LatLongStateDistancesToTraj), size = 1.5) +
  ylab("Lat/Long euclidean distance from Trajectory (nautical miles)") +
  ylim(c(0,20))

latlong_deviation_plot
```

T-test on Latitudes
```{r}
t.test(OwnshipTrajectoryFrame$Latitude, OwnshipStateFrame$LatitudePresent,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

T-Test on Longitudes
```{r}
t.test(OwnshipTrajectoryFrame$Longitude, OwnshipStateFrame$LongitudePresent,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```


## Altitude deviation plots
Unsurprisingly, TrajectoryID = 3 is not too much different in late-flight altitude predictions from TrajectoryID = 1. The rest of the flight state closely follows trajectory. I estimate a regular office building story to be 10 feet in height. This deviation "jitter" magnitude is on the order of 10 feet. Does a plane really oscillate by a story in flight??

I have a feeling that this jitter is due in part to partial Nyquist sampling (1 Hz vs 1/30 Hz). I can try smoothing to correct for this, but the magnitude of the jitter is relatively small compared to the baseline altitude. Would like your input here.

```{r}
altitudeplot<- ggplot() +
  geom_line(data = OwnshipTrajectoryFrame, aes(x = Timestamp, y = Altitude), color = "purple", alpha = 100, size = 1.5) +
  geom_line(data = OwnshipStateFrame, aes(x = Timestamp, y = AltitudeBaroCorrected), color = "cyan", alpha = 100, size = 1.5) +
  ylab("Altitude")

altitudeplot
```

```{r}
altitude_deviation_plot <- ggplot(data = OwnshipStateFrame, aes(x = time, y = deviation)) +
  geom_line(aes(x = Timestamp, y = AltStateDistancesToTraj), size = 1) +
  ylab("Altitude deviation from Trajectory (feet)") 

altitude_deviation_plot
```

Zooming in a little bit to remove the crushing outlier at late timepoints. 7 seconds of state data are masked in this zoom view.

```{r}
altitude_deviation_trunc_plot <- ggplot(data = OwnshipStateFrame, aes(x = time, y = deviation)) +
  geom_line(aes(x = Timestamp, y = AltStateDistancesToTraj), size = 1) +
  ylab("Altitude Euclidean distance from Trajectory (feet)") +
  ylim(c(0,20))

altitude_deviation_trunc_plot
```

Altitude t-test

```{r}
t.test(OwnshipTrajectoryFrame$Altitude, OwnshipStateFrame$AltitudeBaroCorrected,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

## weight deviation

I am a little confused (and disappointed?) about this, but maybe I am expecting too much from Trajectory updates ~90 seconds apart. My expectation, from the example Asha showed earlier this year, was a noticeable improvement in aircraft weight predictions after taking an update. 

Perhaps this will improve further when we have more "flightplan legs" after I correct for the dropped waypoint case.

```{r}
weightplot<- 
  ggplot()+
  geom_line(data = OwnshipTrajectoryFrame, aes(x = Timestamp, y = Weight), size = 1.5, color = "purple")+
  geom_line(data = OwnshipStateFrame, aes(x = Timestamp, y = AircraftWeight), size = 1.5, color = "cyan") +
  ylab("Weight")
weightplot
```

Weight t-test
```{r}
t.test(OwnshipTrajectoryFrame$Weight, OwnshipStateFrame$AircraftWeight,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```



# Definition of State conformance to Trajectory

Here is where I start to get a little uncomfortable. I will try to explain my reasoning for NOT DOING THIS, but I will still show what the results look like with and without thresholding.

So, there are some arbitrary thresholds for an aircraft (State) to be considered conformant to planned route (Trajectory). They are listed below:

Thresholds:

- Latitude, longitude deviation <= 5 NM

- Altitude <= 500 feet

I was asked to filter the data such that State deviation from Trajectory calculations were only done where the aircraft was conforming to the trajectory (within these thresholds).

## Filter State data by lateral and altitude deviations.

```{r}
filtered_OwnshipStateFrame <- OwnshipStateFrame %>%
  filter(LatLongStateDistancesToTraj <= 5 & AltStateDistancesToTraj <= 500)
```

### Post filtration lat, long
```{r}
filtered_flightmap_path <- ggplot()+
 geom_line(data = OwnshipTrajectoryFrame, aes(x = Longitude, y = Latitude), size = 1, color = "purple") +
  geom_point(data = filtered_OwnshipStateFrame, aes(x = LongitudePresent, y = LatitudePresent), size = 0.2, color = "cyan") +
  xlim(c(-90,-82.5))+
  ylim(c(35, 40))

filtered_flightmap_path
```

```{r}
filtered_latlong_deviation_plot <- ggplot(data = filtered_OwnshipStateFrame, aes(x = time, y = deviation)) +
  geom_point(aes(x = Timestamp, y = LatLongStateDistancesToTraj), size = 1.5) +
  ylab("Lat/Long euclidean distance from Trajectory (nautical miles)") +
  ylim(c(0,20))

filtered_latlong_deviation_plot
```

These tests will be flawed because I am comparing all Trajectory data to the filtered State data. Will need to revise this workflow to align Trajectory and State by Timestamp and then do filtration.


T-test on Latitudes
```{r}
t.test(OwnshipTrajectoryFrame$Latitude, filtered_OwnshipStateFrame$LatitudePresent,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

T-Test on Longitudes
```{r}
t.test(OwnshipTrajectoryFrame$Longitude, filtered_OwnshipStateFrame$LongitudePresent,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

### Post-filtration altitude

```{r}
altitudeplot<- ggplot() +
  geom_line(data = OwnshipTrajectoryFrame, aes(x = Timestamp, y = Altitude), color = "purple", alpha = 100, size = 1.5) +
  geom_point(data = filtered_OwnshipStateFrame, aes(x = Timestamp, y = AltitudeBaroCorrected), color = "cyan", alpha = 100, size = 1.5) +
  ylab("Altitude")

altitudeplot
```

```{r}
filtered_altitude_deviation_trunc_plot <- ggplot(data = filtered_OwnshipStateFrame, aes(x = time, y = deviation)) +
  geom_point(aes(x = Timestamp, y = AltStateDistancesToTraj), size = 1) +
  ylab("Altitude Euclidean distance from Trajectory (feet)") +
  ylim(c(0,20))

filtered_altitude_deviation_trunc_plot
```

```{r}
t.test(OwnshipTrajectoryFrame$Altitude, filtered_OwnshipStateFrame$AltitudeBaroCorrected,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

### Post-filtration weight deviation

```{r}
filtered_weightplot<- 
  ggplot()+
  geom_line(data = OwnshipTrajectoryFrame, aes(x = Timestamp, y = Weight), size = 1.5, color = "purple")+
  geom_point(data = filtered_OwnshipStateFrame, aes(x = Timestamp, y = AircraftWeight), size = 1.5, color = "cyan") +
  ylab("Weight")
filtered_weightplot
```


Weight t-test
```{r}
t.test(OwnshipTrajectoryFrame$Weight, filtered_OwnshipStateFrame$AircraftWeight,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95)
```

## an opinion on this

In my naive opinion, this is double correcting for changes in Route. If I am a pilot and ATC tells me to change Route, this workflow will detect that change and take a Trajectory update to account for it. The same thing will be true for TAP-motivated re-routes (at least for the first one). If I deviate from the planned route for any other reason, that should be considered normal flight. Restricting the definition of "normal" is overcorrection.

ALSO, this correction doesn't even really make the prediction look any better. We only keep `r nrow(filtered_OwnshipStateFrame)/nrow(OwnshipStateFrame)*100` % of the data after thresholding. Visually, we don't gain very much in terms of apparent model performance. We also lose interpretability of our t-tests!

I propose revisiting the subject after I correct the flightplan identifier logic, reporting the percentage of conformant datapoints, but not filtering out nonconforming points.

